### 2022-04-16 draft_nb.Rmd
update 4/16 pt1

---
title: "R Notebook"
output:
---
# Intro
The project will be using some macroeconomic knowledge, specifically the Monetarist understanding of economic activity. It's defined by a deceivingly simply formula: ${MV = PQ}$  
According to this definition, M is the money supply, V is the money velocity, P is the price of goods and services, and Q is the quantity of goods and services. Meaning the price of goods can be demonstrated as such: ${P = \frac{MV}{Q}}$  
In terms of economic indicators measured by the Federal Reserve Bank, this formula is best represented by:  
**P - Consumer Price Index for All Urban Consumers: All Items in U.S. City Average (though for this investigation we'll be using the percentage change from previous year figures to represent inflationary movements)**  
**M - Real M2 Money Stock**  
**V - Velocity of M2 Money Stock**  
**Q - Real Gross Domestic Product**  
Luckily, the Fed saved me the trouble of adjusting these figures for CPI myself, and for that I thank them deeply. In terms of a scientific model, there isn't an independent variable we can manipulate freely like in a controlled experiment. However, there is a tool regularly used by the Fed to temper inflation and unemployment, the interest rate. According to macroeconomic theory, increasing interest rates decreases lending/borrowing and thus consumer spending, decreasing inflation at the cost of economic growth. Because it is the only indicator that is so closely related to inflation ***and*** can be directly manipulated --- though often as a response to other economic activity or fiscal policy changes --- it serves as the exposure variable in the following DAG.
```{r}
library(dagitty)
library(ggdag)
library(tidyverse)
library(tidybayes)
library(bsts)
library(rstan)
library(modelr)
library(brms)
library(gganimate)
library(cowplot)
library(ggridges)
library(colorspace)
library(lubridate)
```

```{r}
# NOTE TO RYAN: Let me know if I should have the unobserved variables point towards interest rate as well.
#I know market activity leads to the decision to change the IR, but I don't know if that's the same as influencing it in a way to portray that in the DAG.
#If it is, then all of my variables are affected by unobserved variables which isn't ideal but I can work with it either way.
#And if not, I can include that when I explain my model's weaknesses before continuing.

g <- dagitty::dagitty('
dag {
bb="0,0,1,1"
"Other Indicators/Exogenous Shocks" [latent,pos="0.533,0.067"]
"interest rate" [exposure,pos="0.487,0.541"]
"money supply" [pos="0.595,0.353"]
"money velocity" [pos="0.748,0.309"]
gdp [pos="0.337,0.275"]
inflation [outcome,pos="0.476,0.270"]
"Other Indicators/Exogenous Shocks" -> "money supply"
"Other Indicators/Exogenous Shocks" -> "money velocity"
"Other Indicators/Exogenous Shocks" -> gdp
"Other Indicators/Exogenous Shocks" -> inflation
"interest rate" -> "money supply"
"interest rate" -> gdp
"interest rate" -> inflation
"money supply" -> inflation
"money velocity" -> "interest rate"
"money velocity" -> inflation
gdp -> inflation
}')
ggdag::ggdag_status(g, text = FALSE, use_labels = "name") +
  theme_dag()
```

# Grabbing data from FRED DATA
```{r}
Q = read_csv('gdp.csv', show_col_types = F) %>%
  filter(DATE >= as.Date('1980-01-01'))
P = read_csv('inflation_rate.csv', show_col_types = F) %>%
  filter(DATE >= as.Date('1980-01-01')) %>%
  mutate(CPIAUCSL_PC1 = as.double(CPIAUCSL_PC1)/100)
I = read_csv('interest_rate.csv', show_col_types = F) %>%
  filter(DATE >= as.Date('1980-01-01')) %>%
  mutate(DFF = as.double(DFF)/100)
M = read_csv('money_supply.csv', show_col_types = F) %>%
  filter(DATE >= as.Date('1980-01-01'))
V = read_csv('money_velocity.csv', show_col_types = F) %>%
  filter(DATE >= as.Date('1980-01-01'))
monthly_train <- P %>% inner_join(I, by = 'DATE') %>% inner_join(M, by = 'DATE') %>% left_join(Q, by = 'DATE') %>% left_join(V, by = 'DATE') %>%
  rename(P = CPIAUCSL_PC1, I = DFF, M = M2REAL, Q = GDPC1, V = M2V) %>%
  filter(DATE < as.Date('2015-01-01')) %>%
  mutate(time = 1:n(),
         Q = na.locf(Q),
         V = na.locf(V),
         logitP = qlogis(P),
         m = month(DATE))
monthly_test <- P %>% inner_join(I, by = 'DATE') %>% inner_join(M, by = 'DATE') %>% left_join(Q, by = 'DATE') %>% left_join(V, by = 'DATE') %>%
  rename(P = CPIAUCSL_PC1, I = DFF, M = M2REAL, Q = GDPC1, V = M2V) %>%
  filter(DATE >= as.Date('2015-01-01')) %>%
  mutate(time = 1:n(),
         Q = na.locf(Q),
         V = na.locf(V),
         logitP = qlogis(P),
         m = month(DATE))
# Warning is for April 2022
```

# Using bsts
```{r}
# Bare bones, only seasonal prior, mostly learning way around bsts
model_components <- list()
model_components <- AddSemilocalLinearTrend(model_components, monthly_train$logitP)
fit <- bsts(monthly_train$logitP, model_components, niter = 5000)
plot(fit)
```

# Create Priors about economic indicators and Re-fit
```{r}

```

# Prior Predictive Check - Visualize Priors
```{r}

```

## Iterate, if necessary
```{r}

```

# Create first model - adding other monthly data
```{r}
# Not sure how to add quarterlies yet because of season.duration component in AddSeasonal
# Will need to figure out so I can stratify by Money Velocity
model_components <- list()
model_components <- AddLocalLevel(model_components, y =  monthly_train$logitP)
model_components <- AddSeasonal(model_components, y = monthly_train$logitP, nseasons = 32, season.duration = 4)
#fit2 <- bsts(logitP~I+V, model_components, niter = 5000, data = monthly_train) the bsts works regardless of what's in the model
fit2 <- bsts(monthly_train$logitP, model_components, niter = 5000)
```

```{r}
plot(fit2, 'components')
pred1 <- predict(fit2, horizon = 12, newdata = monthly_test) #this doesn't work when I add predictors due to 'Non-finite response variable' = logit(inflation) being -Inf
plot(pred1)
```

```{r}
forecast_months = 24   # number of months forward to forecast
set.seed(123456)
y_max = .2
y_axis = list(
  coord_cartesian(ylim = c(0, y_max), expand = FALSE),
  scale_y_continuous(labels = scales::percent),
  theme(axis.text.y = element_text(vjust = 0.05))
)
title = labs(x = NULL, y = NULL, subtitle = "US inflation over time")
```

```{r}
fits = monthly_train %>%
  add_draws(plogis(colSums(aperm(fit2$state.contributions, c(2, 1, 3)))))
predictions = monthly_train %$%
  tibble(
    DATE = max(DATE) + months(1:forecast_months),
    m = month(DATE),
    time = max(time) + 1:forecast_months
  ) %>%
  add_draws(plogis(predict(fit2, newdata = monthly_test, horizon = forecast_months, burn = burn)$distribution), value = ".prediction")
predictions_with_last_obs = monthly_train %>% 
  slice(n()) %>% 
  mutate(.draw = list(1:max(predictions$.draw))) %>% 
  unnest() %>% 
  mutate(.prediction = P) %>% 
  bind_rows(predictions)
```

```{r}
#As you can see, these aren't good predictions at all. They cling way too strongly to the priors set around 2%, which I set only because it's the typical benchmark the Fed hopes to hold. I'll be changing the priors.
monthly_train %>%
  ggplot(aes(x = DATE, y = P)) +
  geom_line(aes(y = .value, group = .draw), alpha = 1/20, data = fits %>% sample_draws(100)) +
  geom_line(aes(y = .prediction, group = .draw), alpha = 1/20, data = predictions %>% sample_draws(100)) +
  geom_point()
```

# Convergence Diagnostics (?) or Fake Data Simulation
```{r}

```

# Posterior Sampling
```{r}

```

# Trankplots
```{r}

```

# Stratification for direct and indirect effects
```{r}

```