### 2022-05-07 draft_nb.Rmd
5/6 update

---
title: "R Notebook"
output: html
---
# Intro
The project will be using some macroeconomic knowledge, specifically the Monetarist understanding of economic activity. It's defined by a deceivingly simply formula: ${MV = PQ}$  
According to this definition, M is the money supply, V is the money velocity, P is the price of goods and services, and Q is the quantity of goods and services. Meaning the price of goods can be demonstrated as such: ${P = \frac{MV}{Q}}$  
In terms of economic indicators measured by the Federal Reserve Bank, this formula is best represented by:  
**P - Consumer Price Index for All Urban Consumers: All Items in U.S. City Average (though for this investigation we'll be using the percentage change from previous year figures to represent inflationary movements)**  
**M - Real M2 Money Stock**  
**V - Velocity of M2 Money Stock**  
**Q - Real Gross Domestic Product**  
Luckily, the Fed saved me the trouble of adjusting these figures for CPI myself, and for that I thank them deeply. In terms of a scientific model, there isn't an independent variable we can manipulate freely like in a controlled experiment. However, there is a tool regularly used by the Fed to temper inflation and unemployment, the interest rate. According to macroeconomic theory, increasing interest rates decreases lending/borrowing and thus consumer spending, decreasing inflation at the cost of economic growth. Because it is the only indicator that is so closely related to inflation ***and*** can be directly manipulated --- though often as a response to other economic activity or fiscal policy changes --- it serves as the exposure variable in the following directed acyclic graph (DAG).
```{r, echo=FALSE, warning=FALSE}
library(dagitty)
library(ggdag)
library(tidyverse)
library(tidybayes)
library(bsts)
library(rstan)
library(modelr)
library(brms)
library(gganimate)
library(cowplot)
library(ggridges)
library(colorspace)
library(lubridate)
library(rethinking)
```

```{r, echo=FALSE}
g <- dagitty::dagitty('
dag {
bb="0,0,1,1"
"I-1" [pos="0.080,0.082"]
"IR-1" [exposure,pos="0.078,0.219"]
I [outcome,pos="0.487,0.082"]
IR [pos="0.487,0.223"]
M [pos="0.417,0.416"]
U [latent,pos="0.254,0.539"]
V [pos="0.282,0.340"]
gdp [pos="0.248,0.272"]
"I-1" -> "IR-1"
"I-1" -> I
"IR-1" -> I
"IR-1" -> IR
"IR-1" -> gdp
I -> IR
M -> I
U -> I
U -> M
U -> V
U -> gdp
V -> "IR-1"
V -> I
gdp -> I
}')
ggdag::ggdag_status(g, text = FALSE, use_labels = "name") +
  theme_dag()
```

It is crucial to remember that this model, like all models, is a simplified portrayal of the interactions it is meant to describe. Due to the compounding impacts of exogenous shocks and indicators outside the scope of the model, we cannot model the direct effect of the interest rate on inflation. However, we are able to model the total effect of the interest rate on inflation if we control for money velocity.

The next step will be to create our priors and build a generative model. The loop I crafted to do so has its limitations, though I wasn't able to find any functions or loops to use as reference. The goal is to create time series that behave similarly to the indicators they're meant to mimic.
# Generative Model
## Forming data from priors
```{r, echo=FALSE}
set.seed(9)
N <- 600 # Number of observations
time <- seq(1, N, 1) # Time vector
P <- c(4.5) # "true" mu
for(i in seq(599)){
  p <- rnorm(1, mean = P[i], rnorm(1, 0.3,0.1)) # Makes sure next point stems from previous
  P <- c(P, p)}
P <- round(P/100, 2)
plot(time, P, 'l')

set.seed(18)
IR <- c(1)
for(i in seq(599)){
  ir <- round(rnorm(1, mean = IR[i], 0.05),1)
  IR <- c(IR, ir)}
IR <- IR / 100
plot(time, IR, 'l')

set.seed(4)
V <- c(2)
for(i in seq(599)){
  v <- round(rnorm(1, mean = V[i], 0.04), 2)
  V <- c(V, v)}
plot(time, V, 'l')
```

Here we have fake data for the indicators going into the model. We use those to model the relationship using the priors we set as well as priors included in the bayesian structure time series (bsts) package.

## Prior Predictive Check - Visualize Priors
```{r}
gm <- cbind.data.frame(time, P, IR, V)
gm_test <- gm[501:600,]
gm <- gm[1:500,]
gmod_components <- list()
gmod_components <- AddStudentLocalLinearTrend(gmod_components, y = gm$P)
gmod_components <- AddSeasonal(gmod_components, y = gm$P, nseasons = 50, season.duration = 4)
gmod <- bsts(P~IR+V, gmod_components, niter = 5000, data = gm)

gpred <- predict(gmod, horizon = 10, gm_test)
plot(gpred, ylim= c(-0.1, 0.1))
```

```{r}
forecast_time = 100
fits = gm %>%
  add_draws(colSums(aperm(gmod$state.contributions, c(2, 1, 3))))

predictions = data.frame(time = max(gm$time) + 1:forecast_time) %>%
  add_draws(predict(gmod, newdata = gm_test, horizon = forecast_time)$distribution, value = ".prediction")

predictions_with_last_obs = gm %>% 
  slice(n()) %>% 
  mutate(.draw = list(1:max(predictions$.draw))) %>% 
  unnest(cols = c(.draw)) %>% 
  mutate(.prediction = P) %>% 
  bind_rows(predictions)
gm %>%
  ggplot(aes(x = time, y = P)) +
  geom_line(aes(y = .value, group = .draw), alpha = 1/20, data = fits %>% sample_draws(100)) +
  geom_line(aes(y = .prediction, group = .draw), alpha = 1/20, data = predictions %>% sample_draws(100)) +
  geom_point()
```

We see from the prior predictives that the model and priors are behaving in a way we'd expect. The values themselves aren't what you'd expect from inflation though this a result of struggling to find rnorm values that looked like the inflation curve. We may attribute the fuzziness in predictions through even the observed time values to the fact that the generated priors aren't actually related. Additionally, seeing such a wide range of prior predicted values is expected and satisfactory for something as unpredictable as the economy.

With a working set of priors, we can now move onto modeling real data.

# Modeling with Real Data
```{r}
# Load data
gdp_file = "https://github.com/SamL153/bayesian_stats_project/raw/master/gdp.csv"
inflation_file = "https://github.com/SamL153/bayesian_stats_project/raw/master/inflation_rate.csv"
interest_file = "https://github.com/SamL153/bayesian_stats_project/raw/master/interest_rate.csv"
msupply_file = "https://github.com/SamL153/bayesian_stats_project/raw/master/money_supply.csv"
mvelocity_file = "https://github.com/SamL153/bayesian_stats_project/raw/master/money_velocity.csv"
Q = read_csv(gdp_file, show_col_types = F) %>%
  filter(DATE >= as.POSIXlt('1980-01-01'))
P = read_csv(inflation_file, show_col_types = F) %>%
  filter(DATE >= as.POSIXlt('1980-01-01')) %>%
  mutate(CPIAUCSL_PC1 = as.double(CPIAUCSL_PC1)/100)
I = read_csv(interest_file, show_col_types = F) %>%
  filter(DATE >= as.POSIXlt('1980-01-01')) %>%
  mutate(DFF = as.double(DFF)/100)
M = read_csv(msupply_file, show_col_types = F) %>%
  filter(DATE >= as.POSIXlt('1980-01-01'))
V = read_csv(mvelocity_file, show_col_types = F) %>%
  filter(DATE >= as.POSIXlt('1980-01-01'))
monthly_train <- P %>% inner_join(I, by = 'DATE') %>% inner_join(M, by = 'DATE') %>% left_join(Q, by = 'DATE') %>% left_join(V, by = 'DATE') %>%
  rename(P = CPIAUCSL_PC1, I = DFF, M = M2REAL, Q = GDPC1, V = M2V) %>%
  filter(DATE < as.POSIXlt('2021-07-01')) %>%
  mutate(time = 1:n(),
         Q = na.locf(Q),
         V = na.locf(V),
         m = month(DATE))
monthly_test <- P %>% inner_join(I, by = 'DATE') %>% inner_join(M, by = 'DATE') %>% left_join(Q, by = 'DATE') %>% left_join(V, by = 'DATE') %>%
  rename(P = CPIAUCSL_PC1, I = DFF, M = M2REAL, Q = GDPC1, V = M2V) %>%
  filter(DATE >= as.POSIXlt('2021-07-01')) %>%
  mutate(time = 1:n(),
         Q = na.locf(Q),
         V = na.locf(V),
         m = month(DATE))
I_m1_train <- rbind(I[1,], I) %>%
  filter(DATE < as.POSIXlt('2021-06-01')) %>%
  rename(I_m1 = DFF)
I_m1_test <- I %>%
  filter(DATE >= as.POSIXlt('2021-06-01'), DATE < as.POSIXlt('2022-02-01')) %>%
  rename(I_m1 = DFF)
P_m1_train <- read_csv(inflation_file, show_col_types = F) %>%
  filter(DATE >= as.POSIXlt('1979-12-01'), DATE < as.POSIXlt('2021-06-01')) %>%
  mutate(CPIAUCSL_PC1 = as.double(CPIAUCSL_PC1)/100) %>%
  rename(P_m1 = CPIAUCSL_PC1)
P_m1_test <- (read_csv(inflation_file, show_col_types = F) %>%
  filter(DATE >= as.POSIXlt('2021-06-01'), DATE < as.POSIXlt('2022-02-01')) %>%
  mutate(CPIAUCSL_PC1 = as.double(CPIAUCSL_PC1)/100) %>%
  rename(P_m1 = CPIAUCSL_PC1))
monthly_train <- cbind(monthly_train, P_m1_train[,2], I_m1_train[,2])
monthly_test <- cbind(monthly_test, P_m1_test[,2], I_m1_test[,2])
```

## Create first model - adding other monthly data
```{r}
model_components <- list()
model_components <- AddStudentLocalLinearTrend(model_components, y =  monthly_train$P)
model_components <- AddSeasonal(model_components, y = monthly_train$P, nseasons = 41, season.duration = 4)
model <- bsts(P~I_m1+V, model_components, niter = 10000, data = monthly_train)
```

```{r}
plot(model, 'components')
pred1 <- predict(model, horizon = 8, newdata = monthly_test)
plot(pred1)
```

```{r}
forecast_months = 8 # number of months forward to forecast
set.seed(123456)
y_max = .1
y_axis = list(
  coord_cartesian(ylim = c(-0.02, y_max), expand = FALSE),
  scale_y_continuous(labels = scales::percent),
  theme(axis.text.y = element_text(vjust = 0.05))
)
title = labs(x = NULL, y = NULL, subtitle = "US inflation over time")
```

```{r}
fits = monthly_train %>%
  add_draws(colSums(aperm(model$state.contributions, c(2, 1, 3))))
predictions = monthly_train %$%
  tibble(
    DATE = max(DATE) + months(1:forecast_months),
    m = month(DATE),
    time = max(time) + 1:forecast_months
  ) %>%
  add_draws(predict(model, newdata = monthly_test, horizon = forecast_months)$distribution, value = ".prediction")
predictions_with_last_obs = monthly_train %>% 
  slice(n()) %>% 
  mutate(.draw = list(1:max(predictions$.draw))) %>% 
  unnest(cols = c(.draw)) %>% 
  mutate(.prediction = P) %>% 
  bind_rows(predictions)
```

```{r}
monthly_train %>%
  ggplot(aes(x = DATE, y = P)) +
  geom_line(aes(y = .value, group = .draw), alpha = 1/20, data = fits %>% sample_draws(100)) +
  geom_line(aes(y = .prediction, group = .draw), alpha = 1/20, data = predictions %>% sample_draws(100)) +
  geom_point()
```

```{r}
since_year = 2014
set.seed(123456)
fit_color = "#3573b9"
fit_color_fill = hex(mixcolor(.6, sRGB(1,1,1), hex2RGB(fit_color)))
prediction_color = "#e41a1c"
prediction_color_fill = hex(mixcolor(.6, sRGB(1,1,1), hex2RGB(prediction_color)))

x_axis = list(
  scale_x_date(date_breaks = "1 years", labels = year),
  theme(axis.text.x = element_text(hjust = 0.1))
)
arrow_spec = arrow(length = unit(8, "points"), type = "closed")
fit_annotation = list(
  annotate("text", x = ymd("2014-04-08"), y = .031, hjust = 0, vjust = 0, lineheight = 1.1,
    label = 'Uncertainty in what\ninflation'),
  annotate("text", x = ymd("2014-04-08"), y = .031, hjust = 0, vjust = 0, lineheight = 1.1,
    label = '\n             was', color = fit_color, fontface = "bold"),
  annotate("segment", x = ymd("2015-03-01"), xend = ymd("2015-03-01"), y = .03, yend = .001, linejoin = "mitre",
    color = fit_color, size = 1, arrow = arrow_spec)
)
prediction_annotation = list(
  annotate("text", x = ymd("2020-02-16"), y = .031, hjust = 0, vjust = 0, lineheight = 1.1,
    label = 'Uncertainty in what \ninflation'),
  annotate("text", x = ymd("2020-02-16"), y = .031, hjust = 0, vjust = 0, lineheight = 1.1,
    label = '\n             will be', color = prediction_color, fontface = "bold"),
  annotate("segment", x = ymd("2021-03-01"), xend = ymd("2021-03-01"), y = .03, yend = .02, linejoin = "mitre",
    color = prediction_color, size = 1, arrow = arrow_spec)
)

monthly_train %>%
  filter(year(DATE) >= since_year) %>%
  ggplot(aes(x = DATE, y = P)) +
  geom_line(aes(y = .value, group = .draw), alpha = 1/30, color = fit_color, size = .75, # For each year since 2008, take 100 samples and use those to create lines which are blurred to show what i believe is the posterior predictives for each year.
    data = fits %>% filter(year(DATE) >= since_year) %>% sample_draws(100)) +
  geom_line(aes(y = .prediction, group = .draw), alpha = 1/20, color = prediction_color, size = .75, # Does prior predictives based on fit. Clearly they scatter more
    data = predictions %>% sample_draws(100)) + # point estimates from initial data
  geom_point(size = 0.75) +
  fit_annotation +
  prediction_annotation +
  y_axis +
  x_axis +
  title
```

```{r}
monthly_train %>%
  filter(year(DATE) >= since_year) %>%
  ggplot(aes(x = DATE, y = P)) +
  stat_lineribbon(aes(y = .value), fill = adjustcolor(fit_color, alpha.f = .25), color = fit_color, .width = .95,
    data = fits %>% filter(year(DATE) >= since_year)) +
  stat_lineribbon(aes(y = .prediction), fill = adjustcolor(prediction_color, alpha.f = .25), color = prediction_color, .width = .95,
    data = predictions) +
  geom_line(aes(y= P), data = monthly_test)+
  geom_point(size = 0.75) +
  fit_annotation +
  prediction_annotation +
  y_axis +
  x_axis +
  title
```

```{r}
n_bands = 40

monthly_train %>%
  filter(year(DATE) >= since_year) %>%
  ggplot(aes(x = DATE, y = P)) +
  stat_lineribbon(aes(y = .value), fill = fit_color, alpha = 1/n_bands, .width = ppoints(n_bands), 
    data = fits %>% filter(year(DATE) >= since_year), color = NA) +
  stat_lineribbon(aes(y = .prediction), fill = prediction_color, alpha = 1/n_bands, .width = ppoints(n_bands),
    data = predictions, color = NA) +
  geom_point(size = 0.75) +
  fit_annotation +
  prediction_annotation +
  y_axis +
  x_axis +
  title
```

# Convergence Diagnostics (?) or Fake Data Simulation
```{r}

```

# Posterior Sampling
```{r}

```

# Trankplots
```{r}

```

# Stratification for direct and indirect effects
```{r}

```